### 2015-05-22 reports/paper/introduction.tex
spell check

Accuracte indoor locazation allows creation of noval applications with surrounding awareness that uses position and movment information as input. One application is to allow users to draw with music without physically touching the computer. Another example is to build AI games with physical pieces such as toy car racing where the computer controls some them. In this work, we aim to build such a source localization system that is portable, inexpensive, yet resonably accuracte for localization in a small area.

@@

In this paper, we first discussed in Section II some prior relevant research and approaches in sound localization. In Section III, we evaluated different array achitectures and their impact on localization accuracy. We demonstrated that a two array system outperforms a single array system of similar physical dimension. In Section IV, we presented the chosen architecture along with hardware details. Finally, experiment details and results are presented in Section V.

>>>>

Accurate indoor localization allows creation of novel applications with surrounding awareness that uses position and movement information as input. One application is to allow users to draw with music without physically touching the computer. Another example is to build AI games with physical pieces such as toy car racing where the computer controls some them. In this work, we aim to build such a source localization system that is portable, inexpensive, yet reasonably accurate for localization in a small area.

@@

In this paper, we first discussed in Section II some prior relevant research and approaches in sound localization. In Section III, we evaluated different array architectures and their impact on localization accuracy. We demonstrated that a two array system outperforms a single array system of similar physical dimension. In Section IV, we presented the chosen architecture along with hardware details. Finally, experiment details and results are presented in Section V.

### 2015-05-22 reports/paper/prior.tex
spell check

Cross-correlation has also been used to measure signal arrival time differences\cite{tdoa:mit2, tdoa:micloc, tdoa:3}. Cross-correlation with prefilterings is known as \emph{generalized cross correlation (GCC)}. Different prefilterings have been investigated to improve arrival time difference estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.

Under the GCC framework, the arrival time difference $t_0$ between two signals $x_1(t)$ and $x_2(t)$ can be estimated as:
\begin{eqnarray} \label{eq:gcc}
t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\\label{eq:gcc2}
R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
\end{eqnarray}
, where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transform of $x_1(t)$ and $x_2(t)$. $W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator. We focused on three ways of prefiltering the signal:
\begin{description}[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.
\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divideded by its magnitude. Only phase information contributes to delay estimation.
\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. part of magnitude information is included in delay estimation.
\end{description}

>>>>

Cross-correlation has also been used to measure signal arrival time differences\cite{tdoa:mit2, tdoa:micloc, tdoa:3}. Cross-correlation with prefiltering is known as \emph{generalized cross correlation (GCC)}. Different prefiltering approaches have been investigated to improve arrival time difference estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.

Under the GCC framework, the arrival time difference $t_0$ between two signals $x_1(t)$ and $x_2(t)$ can be estimated as:
\begin{eqnarray} \label{eq:gcc}
t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\\label{eq:gcc2}
R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
\end{eqnarray}
, where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transform of $x_1(t)$ and $x_2(t)$. $W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator. We focused on three ways of prefiltering the signal:
\begin{description}[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.
\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divided by its magnitude. Only phase information contributes to delay estimation.
\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. Part of magnitude information is included in delay estimation.
\end{description}

### 2015-05-22 reports/paper/simulation.tex
spell check

To see how precision affects localization accuracy, we simulated two microphones placed at: $M_1:(x=-10\mbox{ cm},y=0\mbox{ cm})$ and $M_2:(x=10\mbox{ cm},y=0\mbox{ cm})$. A test sound source is emitted at point $P$ which is $50$ centimeters away from the origin $(0,0)$. Let $2a$ denote the difference of distance between $P$ and two microphones:
\[
2a = P M_1 - P M_2 
\]
, then fig~\ref{fig:sim_2_5} shows the region $R$ where all points have difference of distance to two microphones close to $a$ within $1$ cm:
\[
R=\{\hat P: |(\hat P M_1 - \hat P M_2) - 2a|< 1 \mbox{ cm}\}
\]
Intuitively, points in $R$ have difference of distance to two microphones very similar to each other. Looking at fig~\ref{fig:sim_2_5}, we can still see that $R$ has the shape of a hyperbola, but with an uncertainty region around it. The thickness of the uncertainty region is not uniform around the hyperbola, the farther away the point is, the larger the uncertainty region becomes. This indicates for the the same delta distance movement it will generate smaller difference of distance change when the source is farther away from the array. The size of the uncertianty region is also angle dependent: points closer to the line connecting microphones have larger region compared to points close to the line bisecting microphones.

>>>>

To see how precision affects localization accuracy, we simulated two microphones placed at: $M_1:(x=-10\mbox{ cm},y=0\mbox{ cm})$ and $M_2:(x=10\mbox{ cm},y=0\mbox{ cm})$. A test sound source is emitted at point $P$ which is $50$ centimeters away from the origin $(0,0)$. Let $2a$ denote the difference of distance between $P$ and two microphones:
\[
2a = P M_1 - P M_2 
\]
, then fig~\ref{fig:sim_2_5} shows the region $R$ where all points have difference of distance to two microphones close to $a$ within $1$ cm:
\[
R=\{\hat P: |(\hat P M_1 - \hat P M_2) - 2a|< 1 \mbox{ cm}\}
\]
Intuitively, points in $R$ have difference of distance to two microphones very similar to each other. Looking at fig~\ref{fig:sim_2_5}, we can still see that $R$ has the shape of a hyperbola, but with an uncertainty region around it. The thickness of the uncertainty region is not uniform around the hyperbola, the farther away the point is, the larger the uncertainty region becomes. This indicates for the same delta distance movement it will generate smaller difference of distance change when the source is farther away from the array. The size of the uncertainty region is also angle dependent: points closer to the line connecting microphones have larger region compared to points close to the line bisecting microphones.

### 2015-05-22 reports/paper/system.tex
spell check

To handle the uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation output~\ref{eq:gcc2} as a measure of the likelihood of different arrival time differences. For each microphone array, we build a heatmap for the region. The intensity at each point represents the likelihood of it being the source. For each point on the grid, the theoretical TDOA to each microphone pair can be precomputed. Then the heatmap can be generated by going through points on the grid and performing a lookup using equation~\ref{eq:gcc2}. With three microphones $m_1,m_2,$ and $m_3$, there are three microphone pairs: $m_1m_2,m_1m_3,$ and $m_2m_3$. The theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$. Then the liklihood map $L(x,y)$ can be built by superposing the liklihood from each microphone pair:
\begin{eqnarray*}
L(x,y) &=& R_{m_1,m_2}(D_{m_1,m_2}(x,y)) + R_{m_1,m_3}(D_{m_1,m_3}(x,y)) \\
 & & +R_{m_2,m_3}(D_{m_2,m_3}(x,y)) 
\end{eqnarray*}
,where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_3}(\tau)$, and $R_{m_2,m_3}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.

Liklihood maps from two arrays can be combined into the final liklihood map:
\begin{equation}\label{eq:combine_l}
L(x,y) = L_1(x,y) L_2(x,y)
\end{equation}
, where $L_1(x,y)$ and $L_2(x,y)$ represent the liklihood map from array $1$ and array $2$.

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{left.png}
    \caption{localization with only array 1}
    \label{fig:liklihood1}
  \end{subfigure}
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{right.png}
    \caption{localization with only array 2}
    \label{fig:liklihood2}
  \end{subfigure}
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{combined.png}
    \caption{localization with both arrays}
    \label{fig:liklihood3}
  \end{subfigure}
  \caption{Liklihood maps for localizating point at $(0.0,-0.3)$ m}
  \label{fig:liklihood}
\end{figure}

To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$. It shows the individual liklihood map from each array and also the combined liklihood map according to equation~\ref{eq:combine_l}. Individual arrays give accurate angle estimate, but have high uncertainty in distance estimate. The combined liklihood map demonstrated that by merging estimates from two arrays the system is able to perform more accurate localization.

From a timing point of view, the micro-controller spends $12$ millisecond on sampling the microphone data before sending it to a computer for processing. Sending the data through the USB port takes another $15$ millisecond, and processing on the computer takes around $50$ millisecond. Therefore, the totoal time lag between sound source and localization is around $80$ millisecond.

>>>>

To handle the uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation output~\ref{eq:gcc2} as a measure of the likelihood of different arrival time differences. For each microphone array, we build a heatmap for the region. The intensity at each point represents the likelihood of it being the source. For each point on the grid, the theoretical TDOA to each microphone pair can be precomputed. Then the heatmap can be generated by going through points on the grid and performing a lookup using equation~\ref{eq:gcc2}. With three microphones $m_1,m_2,$ and $m_3$, there are three microphone pairs: $m_1m_2,m_1m_3,$ and $m_2m_3$. The theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$. Then the likelihood map $L(x,y)$ can be built by superposing the likelihood from each microphone pair:
\begin{eqnarray*}
L(x,y) &=& R_{m_1,m_2}(D_{m_1,m_2}(x,y)) + R_{m_1,m_3}(D_{m_1,m_3}(x,y)) \\
 & & +R_{m_2,m_3}(D_{m_2,m_3}(x,y)) 
\end{eqnarray*}
,where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_3}(\tau)$, and $R_{m_2,m_3}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.

Likelihood maps from two arrays can be combined into the final likelihood map:
\begin{equation}\label{eq:combine_l}
L(x,y) = L_1(x,y) L_2(x,y)
\end{equation}
, where $L_1(x,y)$ and $L_2(x,y)$ represent the likelihood map from array $1$ and array $2$.

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{left.png}
    \caption{localization with only array 1}
    \label{fig:liklihood1}
  \end{subfigure}
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{right.png}
    \caption{localization with only array 2}
    \label{fig:liklihood2}
  \end{subfigure}
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{combined.png}
    \caption{localization with both arrays}
    \label{fig:liklihood3}
  \end{subfigure}
  \caption{Likelihood maps for localizing point at $(0.0,-0.3)$ m}
  \label{fig:liklihood}
\end{figure}

To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$. It shows the individual likelihood map from each array and also the combined likelihood map according to equation~\ref{eq:combine_l}. Individual arrays give accurate angle estimate, but have high uncertainty in distance estimate. The combined likelihood map demonstrated that by merging estimates from two arrays the system is able to perform more accurate localization.

From a timing point of view, the micro-controller spends $12$ millisecond on sampling the microphone data before sending it to a computer for processing. Sending the data through the USB port takes another $15$ millisecond, and processing on the computer takes around $50$ millisecond. Therefore, the total time lag between sound source and localization is around $80$ millisecond.
